\LARGE{{\bf{Gravitational Wave Signals from Multiple Hidden Sectors} \\ }}

### 2019-09-03 gwnnDraft02.tex
signal updated

%%%%%%%%%%%new definitions: 
\def\lsim{\mathrel{\rlap{\lower4pt\hbox{\hskip1pt$\sim$}}
  \raise1pt\hbox{$<$}}}
\def\gsim{\mathrel{\rlap{\lower4pt\hbox{\hskip1pt$\sim$}}
  \raise1pt\hbox{$>$}}}
\newcommand{\vev}[1]{ \left\langle {#1} \right\rangle }
\newcommand{\bra}[1]{ \langle {#1} | }
\newcommand{\ket}[1]{ | {#1} \rangle }
\newcommand{\ev}{ {\rm eV} }
\newcommand{\kev}{{\rm keV}}
\newcommand{\mev}{{\rm MeV}}
\newcommand{\gev}{{\mathrm GeV}}
\newcommand{\tev}{{\rm TeV}}
\newcommand{\mpl}{$M_{Pl}$}
\newcommand{\mw}{$M_{W}$}
\newcommand{\Ft}{F_{T}}
\newcommand{\Zparity}{\mathbb{Z}_2}
\newcommand{\BLambda}{\boldsymbol{\lambda}}
\newcommand{\met}{\;\not\!\!\!{E}_T}
\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}
\newcommand{\nn}{\nonumber}
\newcommand{\hc}{\mathrm{h.c.}}
\newcommand{\eps}{\epsilon}
\newcommand{\bwt}{\begin{widetext}}
\newcommand{\ewt}{\end{widetext}}
\newcommand{\draftnote}[1]{{\bf\color{blue} #1}}

\newcommand{\cO}{{\cal O}}
\newcommand{\cL}{{\cal L}}
\newcommand{\cM}{{\cal M}}

%References  
\newcommand{\fref}[1]{Fig.~\ref{fig:#1}} 
\newcommand{\eref}[1]{Eq.~\eqref{eq:#1}} 
\newcommand{\aref}[1]{Appendix~\ref{app:#1}}
\newcommand{\sref}[1]{Section~\ref{sec:#1}}
\newcommand{\tref}[1]{Table~\ref{tab:#1}}

%\title{\LARGE{{\bf{Piled up gravitational waves:} \\
%\bf{Searching for new signals of $N$naturalness} }}}
\title{\LARGE{{\bf{Gravitational Wave Signals from Multiple Hidden Sectors} \\
}}}
\author{{\bf {Paul Archer-Smith, Dylan Linthorne, and Daniel Stolarski}}}

\affiliation{
Ottawa-Carleton  Institute  for  Physics,  Carleton  University,\\
1125  Colonel  By  Drive,  Ottawa,  Ontario  K1S  5B6,  Canada
}

\email{
PaulSmith3@cmail.carleton.ca \\
Dylan.linthorne@carleton.ca \\
}

\begin{abstract}
We explore the possibility of detecting gravitational waves generated by 1st order phase transitions in multiple dark sectors. Nnaturalness is taken as a sample model that features multiple additional sectors, many of which undergo phase transitions that produce gravitational waves. We examine the cosmological history of this framework and determine the gravitational wave profiles generated. These profiles are checked against projections of next generation gravitational wave experiments, demonstrating that Nnaturalness can indeed produce unique gravitational wave signatures that will be probed by these future experiments. 
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction} 
\label{sec:intro}

@@

In this setup, there are two qualitatively different kinds of sectors:
\begin{itemize}
\item \textbf{Standard Sectors}: Those with $m_H^2 < 0$ where electroweak symmetry is broken by the vacuum expectation value (VEV) of a fundamental scalar. As in~\cite{Arkani-Hamed:2016rle}, we assume that the standard sector with the smallest absolute value of $m_H^2$ is the Standard Model.
\item \textbf{Exotic Sectors}: Those with  $m_H^2 > 0$. In this case, electroweak symmetry breaking is preserved below the mass of the Higgs, and broken by the confinement of QCD~\cite{Arkani-Hamed:2016rle,Pisarski:1983ms}.
\end{itemize}
Cosmological observations, particularly limits on extra relativistic degrees of freedom at the time of Big Bang Nucleosynthesis and the time of the formation of the cosmic microwave background (CMB)~\cite{Aghanim:2018eyx}, require that most of the energy in the universe is in the Standard Model sector as we will quantify. Therefore, the hidden sectors cannot be in thermal equilibrium at any time, and the physics of reheating must dump energy preferentially in the Standard Model sector. This can be accomplished in multifield inflation models~\cite{} and with the reheaton method~\cite{Arkani-Hamed:2016rle}. We will also explore alternative parameterizations of reheating that satisfy this condition.

@@

\section{Particle Setup}
\label{sec:nn}

@@

We consider the following Lagrangian as in~\cite{Arkani-Hamed:2016rle}:
\begin{equation}
{\cal L} = \sum_{i=-N/2}^{N/2} {\cal L}_i,
\end{equation}
with ${\cal L}_0 = {\cal L}_{\rm SM}$ being the Standard Model Lagrangian, and ${\cal L}_i$ being a copy of the SM Lagrangian with different fields, but with all dimensionless parameters the same. Each of the Lagrangians does contain a dimensionful operator:
\begin{equation}
{\cal L}_i \subset  - \left(m_H^2\right)_i H_i^\dagger H_i  
\end{equation}
where $H_i$ is a Higgs field in each sector, and the mass term is parametrically given by
\begin{equation}\label{eqn:massParam}
\left(m_H^2\right)_i \sim -\frac{\Lambda_H^2}{N}(2i+r),
\end{equation}
where $\Lambda$ is some high scale cutoff, $N$ is the number of sectors, and $r$ mass parameter in the SM in units of $\Lambda_H^2/N$. We view the parameterization of Eq.~(\ref{eqn:massParam}) as a random distribution in theory space up to the cutoff $\Lambda$, therefore this setup solves the hierarchy problem if $r \sim \mathcal{O}(1)$~\cite{Arkani-Hamed:2016rle}\footnote{Constraints require $r$ to be somewhat smaller than 1.} and our sector is the one that that has the smallest absolute value of the Higgs mass parameter. We have taken for simplicity that there equal numbers sectors with positive and negative $m_H^2$, but this assumption does not affect our analysis. This Nnaturalness framework can be generalized: the various sectors can possess a wide range of particle content that can be freely selected by the model builder. The one exception to this is that ``our" sector must consist of the Standard Model.

From the above Lagrangians, the Higgs in sectors with $i \geq 0$ will get a vev given by:
\begin{equation}\label{eqn:vevs}
v^i = \sqrt{-(m_H^2)_i/\lambda_i} \sim \Lambda_H\sqrt{\frac{2i + r}{\lambda N}},
\end{equation}
$\lambda_i$ is the quartic coefficient of the scalar potential and is the same across all sectors, $\lambda_i=\lambda$. This is another way to see how this framework can solve the hierarchy problem: the Higgs vev is parametrically smaller than the cutoff for $N \gg 1$. 
%The parameter $r$ is a way to adjust the tuning of the model: adjusting $r$ away from 1 (which leads to purely uniform spacing) allows us to make a larger gap between our sector and the next one. Within this work, we looked at variants on this sort of tuning including non-uniform distributions and clustered sector effects; these ideas are discussed in Sec. \ref{sec:gw}.   
%
%
%
%
%$N$naturalness is a model framework that looks to solve the hierarchy problem through the introduction of $N$ mutually non-interacting sectors \cite{Arkani-Hamed:2016rle}. The framework itself is very general: the various sectors can possess a wide range of particle content that can be freely selected by the model builder. The one exception to this is that ``our" sector must consist of the Standard Model. For simplicity, the original paper takes all additional new sectors to be copies of the SM (with the same gauge groups and Yukawa structure); however the Higgs mass parameters of each sector are allowed to taken on values distributed between $-\Lambda_H^2$ and $\Lambda^2_H$, with $\Lambda_H$ being the scale that cuts off quadratic divergences.
%
%This construction leads to sectors that are automatically accidentally tuned to the $1/N$ level, leading to a sector with a mass parameter $m_H^2 \sim \Lambda_H^2/N$. The sector with the smallest non-zero vacuum expectation value (vev) as ``our" sector --- the SM sector. It should be noted that the values of $m_H^2$ pass through zero, resulting in effectively 2 types of sectors: ``standard" sectors like our own with a negative Higgs mass parameter squared, possess a vev, and exhibit electroweak symmetry breaking and ``exotic" sectors with a positive Higgs mass parameter squared that thus feature no electroweak symmetry breaking and a zero vev. 
%
%Keeping with the original convention, we write our mass parameter as
%\begin{equation}\label{eqn:massParam}
%\left(m_H^2\right)_i = -\frac{\Lambda_H^2}{N}(2i+r),
%\end{equation}
%where $i$ denotes the sector ($-\frac{N}{2}\leq i \leq \frac{N}{2}$). $i = 0$ is the sector that possess the smallest vev and is thus ``our" SM sector. \textcolor{green}{Write an equation that shows how the vev scales with $i$.} The vev for each standard sector is given explicitly by
%\begin{equation}\label{eqn:vevs}
%v^i = m^i_H/\sqrt{\lambda_i} = \sqrt{\frac{2i + r}{N}}\left(\frac{\Lambda_H}{\lambda}\right),
%\end{equation}
%$\lambda_i$ is the quartic coefficient of the scalar potential and is the same across all sectors. The parameter $r$ is a way to adjust the tuning of the model: adjusting $r$ away from 1 (which leads to purely uniform spacing) allows us to make a larger gap between our sector and the next one. Within this work, we looked at variants on this sort of tuning including non-uniform distributions and clustered sector effects; these ideas are discussed in Sec. \ref{sec:gw}.   
%
%As mentioned above, half of all sectors in this model feature a Higgs mass parameter such that $m_H^2 < 0$. 
The ``standard sectors'' with $i>0$ feature electroweak symmetry breaking just like in the SM, however the vevs produced scale with the changing mass parameter: $v_i \sim v_{\rm SM}\sqrt{i}$. As the masses scale with the vev, this leads to the masses of the $W$, $Z$ and fermions within each sector increasing with $i$. The consequences of this on the confinement scale of QCD in the $i \geq 1$ sectors is further discussed in Sec. \ref{sec:dQCD}.

The ``exotic sectors'' with $i<0$ provide a radical departure from our own. $m_H^2 > 0$ leads to no vev for the Higgs, and electroweak symmetry is only broken at very low scales due to the phase transition from free quarks to confinement at the QCD scale $\Lambda_{QCD}$~\cite{}, and the masses of the $W$ and $Z$ are comparable to those of QCD resonances. The mass of fundamental fermions are produced via four-fermion interactions generated after integrating out the SU$(2)$ Higgs multiplet. This leads to very light fermions: 
\begin{equation}\label{eqn:exoticFM}
m_f \sim y_f y_t \Lambda_{QCD}^3/(m^2_H)_i \leq 100 \,\ev,
\end{equation}
with $y_f$ representing the Yukawa coupling to fermion $f$. As we will see, the extremely light quarks that appear in these sectors dramatically change the nature of the QCD phase transition --- unlike the SM, the transition is strongly first order. Again, this is further developed in Sec. \ref{sec:dQCD}. Crucially, this results in the production of gravitational waves. This is the physical signature we're interested in exploring within this paper; the calculation and results are presented in Sec. \ref{sec:gw}.

\section{QCD Phase Transition}
\label{sec:dQCD}
%
%\begin{figure*}[tb]
%\centering
%\begin{minipage}[c]{\textwidth}
%\includegraphics[width=1.0\textwidth]{columbia.png}
%\end{minipage}
%\hfill
%\caption{Two possible scenarios for the Columbia Plot with the up and down quark mass taken to be identical. The ``physical point" is the Standard Model. Plot from \cite{Cuteri:2017zcb}
%}
%\label{fig:columbia}
%\end{figure*}

We now study the nature of the QCD phase transition across the different sectors. Due to the confining nature of QCD, the exact nature of the phase transition is often difficult to ascertain analytically and requires the study of lattice simulations. In the SM, it is known that the phase transition is a crossover and does lead to gravitational wave signals~\cite{Aoki:2006we,Bhattacharya:2014ara}. In the general case with 3 or more colours, the phase transition can be strongly first order in two regimes~\cite{Panero:2009tv,SVETITSKY1982423,Pisarski:1983ms}:
\begin{itemize}
\item Three or more light flavours.
\item No light flavours. 
\end{itemize}
Light means mass small compared to the confinement scale $\Lambda_{QCD}$, but what that means quantitatively is not precisely determined. In the SM, the up and down quarks are light, but the strange is not sufficiently light for an SFOPT.

% The results of these studies on the nature of QCD phase transitions is summarized in the so-called ``Columbia" plot (Fig. \ref{fig:columbia}). As seen in the above figure (and explicitly demonstrated through lattice studies \cite{lattice}), our sector is solidly in the weak-cross over area. In order to observe the strong first order transition required for gravitational wave production, at least some of the additional sectors present must feature quarks in either the heavy quark (pure Yang-Mills) or massless limits.    
%
%In the Yang-Mills limit we see $m_{u,d,s} \rightarrow \infty$. This leads to the presence of a global $Z_3$ centre symmetry that is broken at high temperatures but restored low temperatures. Ultimately, this restoration results in a first order phase transition \cite{SVETITSKY1982423}. On the other end of the spectrum, ``massless" quark theories ($m_q \ll \Lambda_{QCD}$ for all quarks) also contain strong phase transitions due to the breakdown of SU$(N)\,\times$ SU$(N)$ chiral symmetry \cite{Pisarski:1983ms}. Both the Yang-Mills and massless limit arguments can be generalized to SU$(N \geq 3)$: coupling both with the additional restriction that we avoid the conformal window indicates that strong first order phase transitions can occur if we have $n = 0$ (pure Yang-Mills) or $3 \leq n < 4N$ (light quark limit) very light flavours.

For the standard sectors in our setup, the quark masses increase with increasing vev, so for sufficiently large $i$, all the quarks will be heavier than $\Lambda_{QCD}$,\footnote{$\Lambda_{QCD}$ does vary with $i$, but the sensitivity is very weak as we will see below.} and those large $i$ sectors will undergo an SFOPT if they are reheated above the the confinement scale. 
%For $N$naturalness (or other mirror sector models), standard sectors with higher vevs possess more massive quarks; eventually, sectors with large enough vevs push us into the upper-right corner. 
Conversely, exotic sectors with zero vev feature six very light quarks, so \textit{all} the exotic sectors undergo SFOPT at the temperature of QCD confinement.

We now calculate the QCD confinement scale for each sector following the same procedure as \cite{Cui:2011wk}. First, due to the parameters of each sector being taken to be identical save for the higgs mass squared (thus $v \neq v_i$, where $v$ is the SM vev), we assume that the strong coupling of every sector is identical at some high scale. Using the one-loop running, the $\beta$ function can be solved:
\begin{equation}\label{eqn:QCDrunningi}
\alpha_{s}^i (\mu) = \frac{2\pi}{11-\frac{2n^i_f}{3}}\frac{1}{\ln{\mu/\Lambda^i}},
\end{equation}
where $n_f^i$ is the number of quark flavours with mass less than $\mu/2$ and $\Lambda^i$ is the scale where it would confine if all quarks remain massless. In the SM defined at scales well above all the quark masses, we have $\Lambda_{QCD} = 89 \pm 5$ in $\overline{MS}$~\cite{PhysRevD.98.030001}.

Because we've set the strong couplings equal at high scales, $\Lambda = \Lambda^i$ for all $i$ at high scales for all sectors. However, since the masses of the quarks in each sector are different, we end up with a unique running of the coupling for each sector. At every quark mass threshold for a given sector, we match the coupling strengths above and below the threshold and determine the new $\Lambda^i$ for the lower scale. For example, at the mass of the top quark, we match a five flavour coupling with the six flavour one:
\begin{equation}
\alpha_s^{i(5)}(2 m^i_t) = \alpha_s^{i(6)}(2 m^i_t) 
\end{equation} 
and thus
\begin{equation}
\Lambda_{(5)}^i = (m_t^{i})^{2/23}(\Lambda_{(6)}^i)^{21/23}.
\end{equation}

Suppressing the $i$'s for notational cleanliness, we can arrive at similar relations at the bottom and charm thresholds
\begin{equation}
\begin{split}
\Lambda_{(4)} = (m_b)^{2/25}(\Lambda_{(5)})^{23/25},
\\
\Lambda_{(3)} = (m_c)^{2/23}(\Lambda_{(4)})^{25/27}.
\end{split}
\end{equation}
These can be combined to show that
\begin{equation}
\Lambda_{(3)} = (m_t m_b m_c)^{2/27}(\Lambda_{(6)})^{21/27}.
\end{equation}
This type of matching procedure can be done as many times as necessary for a given sector. The process terminates when $\Lambda_i$ for a given scale is larger than the next quark mass threshold (i.e running the scale down arrives at the $\Lambda_{QCD}$ phase transition before reaching the next quark mass scale).

@@

The appropriate Lagrangian for a scalar reheaton $\phi$ is: 
\begin{equation}\label{eqn:nnLagrangian}
\cL_\phi \supset -a\phi\sum_i\vert H_i\vert^2 - \frac{1}{2}m^2_\phi \phi^2.
\end{equation}
Note that cross-quartic couplings of the form $\kappa\vert H_i\vert^2\vert H_j\vert^2$ that could potentially ruin the spectrum of $N$naturalness are absent, taken to be suppressed by a very small coupling. Effective Lagrangians for the two different types of sectors present in this theory can be obtained by integrating out of the Higgs bosons in every sector:
\begin{equation}\label{eqn:nnEffLag}
\begin{split}
\cL_\phi^{v \neq 0} &\supset C_1 a y_q\frac{v}{m_h^2}\phi q q^c,
\\
\cL_\phi^{v = 0} &\supset C_2 a \frac{g^2}{16\pi^2 m_H^2}\phi W_{\mu\nu}W^{\mu\nu},
\end{split}
\end{equation}
with $C_i$ representing numerical coefficients, $g$ the weak coupling constant, and $W^{\mu\nu}$ the SU$(2)$ field strength tensor.

>>>>

%%%%%%%%%%%new definitions: 
\def\lsim{\mathrel{\rlap{\lower4pt\hbox{\hskip1pt$\sim$}}
  \raise1pt\hbox{$<$}}}
\def\gsim{\mathrel{\rlap{\lower4pt\hbox{\hskip1pt$\sim$}}
  \raise1pt\hbox{$>$}}}
\newcommand{\vev}[1]{ \left\langle {#1} \right\rangle }
\newcommand{\bra}[1]{ \langle {#1} | }
\newcommand{\ket}[1]{ | {#1} \rangle }
\newcommand{\ev}{ {\rm eV} }
\newcommand{\kev}{{\rm keV}}
\newcommand{\mev}{{\rm MeV}}
\newcommand{\gev}{{\mathrm GeV}}
\newcommand{\tev}{{\rm TeV}}
\newcommand{\mpl}{$M_{Pl}$}
\newcommand{\mw}{$M_{W}$}
\newcommand{\Ft}{F_{T}}
\newcommand{\Zparity}{\mathbb{Z}_2}
\newcommand{\BLambda}{\boldsymbol{\lambda}}
\newcommand{\met}{\;\not\!\!\!{E}_T}
\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}
\newcommand{\nn}{\nonumber}
\newcommand{\hc}{\mathrm{h.c.}}
\newcommand{\eps}{\epsilon}
\newcommand{\bwt}{\begin{widetext}}
\newcommand{\ewt}{\end{widetext}}
\newcommand{\draftnote}[1]{{\bf\color{blue} #1}}
\newcommand{\cO}{{\cal O}}
\newcommand{\cL}{{\cal L}}
\newcommand{\cM}{{\cal M}}
%References  
\newcommand{\fref}[1]{Fig.~\ref{fig:#1}} 
\newcommand{\eref}[1]{Eq.~\eqref{eq:#1}} 
\newcommand{\aref}[1]{Appendix~\ref{app:#1}}
\newcommand{\sref}[1]{Section~\ref{sec:#1}}
\newcommand{\tref}[1]{Table~\ref{tab:#1}}
%\title{\LARGE{{\bf{Piled up gravitational waves:} \\
%\bf{Searching for new signals of $N$naturalness} }}}
\title{\LARGE{{\bf{Gravitational Wave Signals from Multiple Hidden Sectors} \\
}}}
\author{{\bf {Paul Archer-Smith, Dylan Linthorne, and Daniel Stolarski}}}
\affiliation{
Ottawa-Carleton  Institute  for  Physics,  Carleton  University,\\
1125  Colonel  By  Drive,  Ottawa,  Ontario  K1S  5B6,  Canada
}
\email{
PaulSmith3@cmail.carleton.ca \\
Dylan.linthorne@carleton.ca \\
stolar@physics.carleton.ca
}
\begin{abstract}
We explore the possibility of detecting gravitational waves generated by 1st order phase transitions in multiple dark sectors. Nnaturalness is taken as a sample model that features multiple additional sectors, many of which undergo phase transitions that produce gravitational waves. We examine the cosmological history of this framework and determine the gravitational wave profiles generated. These profiles are checked against projections of next generation gravitational wave experiments, demonstrating that Nnaturalness can indeed produce unique gravitational wave signatures that will be probed by these future experiments. 
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction} 
\label{sec:intro} 

@@

In this setup, there are two qualitatively different kinds of sectors:
\begin{itemize}
\item \textbf{Standard Sectors}: Those with $m_H^2 < 0$ where electroweak symmetry is broken by the vacuum expectation value (VEV) of a fundamental scalar. As in~\cite{Arkani-Hamed:2016rle}, we assume that the standard sector with the smallest absolute value of $m_H^2$ is the Standard Model.
\item \textbf{Exotic Sectors}: Those with  $m_H^2 > 0$. In this case, electroweak symmetry breaking is preserved below the mass of the Higgs, and broken by the confinement of QCD~\cite{Arkani-Hamed:2016rle,Pisarski:1983ms}.
\end{itemize}
Cosmological observations, particularly limits on extra relativistic degrees of freedom at the time of Big Bang Nucleosynthesis and the time of the formation of the cosmic microwave background (CMB)~\cite{Aghanim:2018eyx}, require that most of the energy in the universe is in the Standard Model sector as we will quantify. Therefore, the hidden sectors cannot be in thermal equilibrium at any time, and the physics of reheating must dump energy preferentially in the Standard Model sector. This can be accomplished in multifield inflation models~\cite{} and with the reheaton method~\cite{Arkani-Hamed:2016rle}. We will also explore alternative parameterizations of reheating that satisfy this condition.

@@

\section{Particle Setup}
\label{sec:nn}

@@

We consider the following Lagrangian as in~\cite{Arkani-Hamed:2016rle}:
\begin{equation}
{\cal L} = \sum_{i=-N/2}^{N/2} {\cal L}_i,
\end{equation}
with ${\cal L}_0 = {\cal L}_{\rm SM}$ being the Standard Model Lagrangian, and ${\cal L}_i$ being a copy of the SM Lagrangian with different fields, but with all dimensionless parameters the same. Each of the Lagrangians does contain a dimensionful operator:
\begin{equation}
{\cal L}_i \subset  - \left(m_H^2\right)_i H_i^\dagger H_i  
\end{equation}
where $H_i$ is a Higgs field in each sector, and the mass term is parametrically given by
\begin{equation}\label{eqn:massParam}
\left(m_H^2\right)_i \sim -\frac{\Lambda_H^2}{N}(2i+r),
\end{equation}
where $\Lambda$ is some high scale cutoff, $N$ is the number of sectors, and $r$ mass parameter in the SM in units of $\Lambda_H^2/N$. We view the parameterization of Eq.~(\ref{eqn:massParam}) as a random distribution in theory space up to the cutoff $\Lambda$, therefore this setup solves the hierarchy problem if $r \sim \mathcal{O}(1)$~\cite{Arkani-Hamed:2016rle}\footnote{Constraints require $r$ to be somewhat smaller than 1.} and our sector is the one that that has the smallest absolute value of the Higgs mass parameter. We have taken for simplicity that there equal numbers sectors with positive and negative $m_H^2$, but this assumption does not affect our analysis. This Nnaturalness framework can be generalized: the various sectors can possess a wide range of particle content that can be freely selected by the model builder. The one exception to this is that ``our" sector must consist of the Standard Model.
From the above Lagrangians, the Higgs in sectors with $i \geq 0$ will get a vev given by:
\begin{equation}\label{eqn:vevs}
v^i = \sqrt{-(m_H^2)_i/\lambda_i} \sim \Lambda_H\sqrt{\frac{2i + r}{\lambda N}},
\end{equation}
$\lambda_i$ is the quartic coefficient of the scalar potential and is the same across all sectors, $\lambda_i=\lambda$. This is another way to see how this framework can solve the hierarchy problem: the Higgs vev is parametrically smaller than the cutoff for $N \gg 1$. 
%The parameter $r$ is a way to adjust the tuning of the model: adjusting $r$ away from 1 (which leads to purely uniform spacing) allows us to make a larger gap between our sector and the next one. Within this work, we looked at variants on this sort of tuning including non-uniform distributions and clustered sector effects; these ideas are discussed in Sec. \ref{sec:gw}.   
%
%
%
%
%$N$naturalness is a model framework that looks to solve the hierarchy problem through the introduction of $N$ mutually non-interacting sectors \cite{Arkani-Hamed:2016rle}. The framework itself is very general: the various sectors can possess a wide range of particle content that can be freely selected by the model builder. The one exception to this is that ``our" sector must consist of the Standard Model. For simplicity, the original paper takes all additional new sectors to be copies of the SM (with the same gauge groups and Yukawa structure); however the Higgs mass parameters of each sector are allowed to taken on values distributed between $-\Lambda_H^2$ and $\Lambda^2_H$, with $\Lambda_H$ being the scale that cuts off quadratic divergences.
%
%This construction leads to sectors that are automatically accidentally tuned to the $1/N$ level, leading to a sector with a mass parameter $m_H^2 \sim \Lambda_H^2/N$. The sector with the smallest non-zero vacuum expectation value (vev) as ``our" sector --- the SM sector. It should be noted that the values of $m_H^2$ pass through zero, resulting in effectively 2 types of sectors: ``standard" sectors like our own with a negative Higgs mass parameter squared, possess a vev, and exhibit electroweak symmetry breaking and ``exotic" sectors with a positive Higgs mass parameter squared that thus feature no electroweak symmetry breaking and a zero vev. 
%
%Keeping with the original convention, we write our mass parameter as
%\begin{equation}\label{eqn:massParam}
%\left(m_H^2\right)_i = -\frac{\Lambda_H^2}{N}(2i+r),
%\end{equation}
%where $i$ denotes the sector ($-\frac{N}{2}\leq i \leq \frac{N}{2}$). $i = 0$ is the sector that possess the smallest vev and is thus ``our" SM sector. \textcolor{green}{Write an equation that shows how the vev scales with $i$.} The vev for each standard sector is given explicitly by
%\begin{equation}\label{eqn:vevs}
%v^i = m^i_H/\sqrt{\lambda_i} = \sqrt{\frac{2i + r}{N}}\left(\frac{\Lambda_H}{\lambda}\right),
%\end{equation}
%$\lambda_i$ is the quartic coefficient of the scalar potential and is the same across all sectors. The parameter $r$ is a way to adjust the tuning of the model: adjusting $r$ away from 1 (which leads to purely uniform spacing) allows us to make a larger gap between our sector and the next one. Within this work, we looked at variants on this sort of tuning including non-uniform distributions and clustered sector effects; these ideas are discussed in Sec. \ref{sec:gw}.   
%
%As mentioned above, half of all sectors in this model feature a Higgs mass parameter such that $m_H^2 < 0$. 
The ``standard sectors'' with $i>0$ feature electroweak symmetry breaking just like in the SM, however the vevs produced scale with the changing mass parameter: $v_i \sim v_{\rm SM}\sqrt{i}$. As the masses scale with the vev, this leads to the masses of the $W$, $Z$ and fermions within each sector increasing with $i$. The consequences of this on the confinement scale of QCD in the $i \geq 1$ sectors is further discussed in Sec. \ref{sec:dQCD}. 
The ``exotic sectors'' with $i<0$ provide a radical departure from our own. $m_H^2 > 0$ leads to no vev for the Higgs, and electroweak symmetry is only broken at very low scales due to the phase transition from free quarks to confinement at the QCD scale $\Lambda_{QCD}$~\cite{}, and the masses of the $W$ and $Z$ are comparable to those of QCD resonances. The mass of fundamental fermions are produced via four-fermion interactions generated after integrating out the SU$(2)$ Higgs multiplet. This leads to very light fermions: 
\begin{equation}\label{eqn:exoticFM}
m_f \sim y_f y_t \Lambda_{QCD}^3/(m^2_H)_i \leq 100 \,\ev,
\end{equation}
with $y_f$ representing the Yukawa coupling to fermion $f$. As we will see, the extremely light quarks that appear in these sectors dramatically change the nature of the QCD phase transition --- unlike the SM, the transition is strongly first order. Again, this is further developed in Sec. \ref{sec:dQCD}. Crucially, this results in the production of gravitational waves. This is the physical signature we're interested in exploring within this paper; the calculation and results are presented in Sec. \ref{sec:gw}. 
\section{QCD Phase Transition}
\label{sec:dQCD}
%
%\begin{figure*}[tb]
%\centering
%\begin{minipage}[c]{\textwidth}
%\includegraphics[width=1.0\textwidth]{columbia.png}
%\end{minipage}
%\hfill
%\caption{Two possible scenarios for the Columbia Plot with the up and down quark mass taken to be identical. The ``physical point" is the Standard Model. Plot from \cite{Cuteri:2017zcb}
%}
%\label{fig:columbia}
%\end{figure*}
We now study the nature of the QCD phase transition across the different sectors. Due to the confining nature of QCD, the exact nature of the phase transition is often difficult to ascertain analytically and requires the study of lattice simulations. In the SM, it is known that the phase transition is a crossover and does lead to gravitational wave signals~\cite{Aoki:2006we,Bhattacharya:2014ara}. In the general case with 3 or more colours, the phase transition can be strongly first order in two regimes~\cite{Panero:2009tv,SVETITSKY1982423,Pisarski:1983ms}:
\begin{itemize}
\item Three or more light flavours.
\item No light flavours. 
\end{itemize}
Light means mass small compared to the confinement scale $\Lambda_{QCD}$, but what that means quantitatively is not precisely determined. In the SM, the up and down quarks are light, but the strange is not sufficiently light for an SFOPT.
% The results of these studies on the nature of QCD phase transitions is summarized in the so-called ``Columbia" plot (Fig. \ref{fig:columbia}). As seen in the above figure (and explicitly demonstrated through lattice studies \cite{lattice}), our sector is solidly in the weak-cross over area. In order to observe the strong first order transition required for gravitational wave production, at least some of the additional sectors present must feature quarks in either the heavy quark (pure Yang-Mills) or massless limits.    
%
%In the Yang-Mills limit we see $m_{u,d,s} \rightarrow \infty$. This leads to the presence of a global $Z_3$ centre symmetry that is broken at high temperatures but restored low temperatures. Ultimately, this restoration results in a first order phase transition \cite{SVETITSKY1982423}. On the other end of the spectrum, ``massless" quark theories ($m_q \ll \Lambda_{QCD}$ for all quarks) also contain strong phase transitions due to the breakdown of SU$(N)\,\times$ SU$(N)$ chiral symmetry \cite{Pisarski:1983ms}. Both the Yang-Mills and massless limit arguments can be generalized to SU$(N \geq 3)$: coupling both with the additional restriction that we avoid the conformal window indicates that strong first order phase transitions can occur if we have $n = 0$ (pure Yang-Mills) or $3 \leq n < 4N$ (light quark limit) very light flavours.
For the standard sectors in our setup, the quark masses increase with increasing vev, so for sufficiently large $i$, all the quarks will be heavier than $\Lambda_{QCD}$,\footnote{$\Lambda_{QCD}$ does vary with $i$, but the sensitivity is very weak as we will see below.} and those large $i$ sectors will undergo an SFOPT if they are reheated above the the confinement scale. 
%For $N$naturalness (or other mirror sector models), standard sectors with higher vevs possess more massive quarks; eventually, sectors with large enough vevs push us into the upper-right corner. 
Conversely, exotic sectors with zero vev feature six very light quarks, so \textit{all} the exotic sectors undergo SFOPT at the temperature of QCD confinement. 
We now calculate the QCD confinement scale for each sector following the same procedure as \cite{Cui:2011wk}. First, due to the parameters of each sector being taken to be identical save for the higgs mass squared (thus $v \neq v_i$, where $v$ is the SM vev), we assume that the strong coupling of every sector is identical at some high scale. Using the one-loop running, the $\beta$ function can be solved:
\begin{equation}\label{eqn:QCDrunningi}
\alpha_{s}^i (\mu) = \frac{2\pi}{11-\frac{2n^i_f}{3}}\frac{1}{\ln{\mu/\Lambda^i}},
\end{equation}
where $n_f^i$ is the number of quark flavours with mass less than $\mu/2$ and $\Lambda^i$ is the scale where it would confine if all quarks remain massless. In the SM defined at scales well above all the quark masses, we have $\Lambda_{QCD} = 89 \pm 5$ in $\overline{MS}$~\cite{PhysRevD.98.030001}. 
Because we've set the strong couplings equal at high scales, $\Lambda = \Lambda^i$ for all $i$ at high scales for all sectors. However, since the masses of the quarks in each sector are different, we end up with a unique running of the coupling for each sector. At every quark mass threshold for a given sector, we match the coupling strengths above and below the threshold and determine the new $\Lambda^i$ for the lower scale. For example, at the mass of the top quark, we match a five flavour coupling with the six flavour one:
\begin{equation}
\alpha_s^{i(5)}(2 m^i_t) = \alpha_s^{i(6)}(2 m^i_t) 
\end{equation} 
and thus
\begin{equation}
\Lambda_{(5)}^i = (m_t^{i})^{2/23}(\Lambda_{(6)}^i)^{21/23}.
\end{equation}
Suppressing the $i$'s for notational cleanliness, we can arrive at similar relations at the bottom and charm thresholds
\begin{equation}
\begin{split}
\Lambda_{(4)} = (m_b)^{2/25}(\Lambda_{(5)})^{23/25},
\\
\Lambda_{(3)} = (m_c)^{2/23}(\Lambda_{(4)})^{25/27}.
\end{split}
\end{equation}
These can be combined to show that
\begin{equation}
\Lambda_{(3)} = (m_t m_b m_c)^{2/27}(\Lambda_{(6)})^{21/27}.
\end{equation}
This type of matching procedure can be done as many times as necessary for a given sector. The process terminates when $\Lambda_i$ for a given scale is larger than the next quark mass threshold (i.e running the scale down arrives at the $\Lambda_{QCD}$ phase transition before reaching the next quark mass scale).

@@

The appropriate Lagrangian for a scalar reheaton $\phi$ is: 
\begin{equation}\label{eqn:nnLagrangian}
\cL_\phi \supset -a\phi\sum_i\vert H_i\vert^2 - \frac{1}{2}m^2_\phi \phi^2.
\end{equation}
Note that cross-quartic couplings of the form $\kappa\vert H_i\vert^2\vert H_j\vert^2$ that could potentially ruin the spectrum of $N$naturalness are absent, taken to be suppressed by a very small coupling. Effective Lagrangians for the two different types of sectors present in this theory can be obtained by integrating out of the Higgs bosons in every sector:
\begin{equation}\label{eqn:nnEffLag}
\begin{split}
\cL_\phi^{v \neq 0} &\supset C_1 a y_q\frac{v}{m_h^2}\phi q q^c,
\\
\cL_\phi^{v = 0} &\supset C_2 a \frac{g^2}{16\pi^2 m_H^2}\phi W_{\mu\nu}W^{\mu\nu},
\end{split}
\end{equation}
with $C_i$ representing numerical coefficients, $g$ the weak coupling constant, and $W^{\mu\nu}$ the SU$(2)$ field strength tensor.
